---
title: "Navigation and Authentication"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Navigation and Authentication}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates two advanced use cases of web scraping using rvest: Navigation and Authentication. First, you'll learn how to use `rvest` to navigate across web pages. Then, you'll see how you can navigate across web sites that require authentication.

```{r}
library(rvest)
```

## Navigation

In the web scraping context, navigation is the ability to go from page to page in a website. Usually, when using `rvest` to fetch contents from a single web page, you use `read_html()`. However, if you need to navigate from one page to another, `rvest` provides `session()`.

```{r}
turtles <- session("https://www.scrapethissite.com/pages/frames/")
```

`session()` creates a persistent web session that you can navigate in with the following rvest functions `session_jump_to()`, `session_follow_link()`, `session_back()` and `session_forward()`.

So, if you need to check the Carettochelyidae family of Turtles, you can use an absolute url:

```{r}
turtles %>%
  session_jump_to("https://www.scrapethissite.com/pages/frames/?frame=i&family=Carettochelyidae") %>%
  session_history()
```

And then go to the Cheloniidae within the same session (using a relative url in this case):
```{r}
turtles %>%
  session_jump_to("?frame=i&family=Cheloniidae") %>%
  session_history()
```

If you need to follow a link within the web page you can use `session_follow_link()` specifying the name of the link:
```{r}
turtles %>%
  session_jump_to("?frame=i&family=Cheloniidae") %>%
  session_follow_link("Back to all Turtles")
```

You can also use `session_back()` and `session_forward()` to navigate back and forth across pages:
```{r}
turtles %>%
  session_jump_to("?frame=i&family=Cheloniidae") %>%
  session_back() %>%
  session_forward() %>%
  session_history()
```

### Forms

Let's suppose you're fan of the NHL team Boston Bruins and want to retrieve all the campaigns from 1990 to 2011.
In this case, you can submit a form with the text "Boston" and you'll get the expected result, using `html_form()` to extract the form, `html_form_set()` to set the form values and finally submit with `session_submit()`.
```{r}
teams <- session("https://www.scrapethissite.com/pages/forms/")

search <- teams %>%
  html_node("form") %>%
  html_form() %>%
  html_form_set(q = "Boston")

boston_campaigns <- teams %>%
  session_submit(search) %>%
  html_table()

boston_campaigns
```

### rvest<1.0.0

Prior to `rvest 1.0.0` the functions `session()`, `session_jump_to()`, `session_follow_link()`, `session_back()` and `session_forward()` were called `html_session`, `jump_to()`, `follow_link()`, `back()` and `forward()`, so if you encounter any of these in some R code, make sure you upgrade to the latest version of `rvest`.

## Authentication

Most of the web sites rely on forms for login. In this case, you can use the same approach described on the navigation section to provide credentials and log in.

```{r}
gh_session <- session("https://github.com/login")

login <- gh_session %>%
  html_element("form")%>%
  html_form()%>%
  html_form_set(login = "YourGitHubUsername", password = "SuperSecureP@ssw0rd")

github <- gh_session %>%
  session_submit(login, submit = "commit") %>%
  read_html()
github
```
