---
title: "SelectorGadget"
author: "Hadley Wickham"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Selectorgadget}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
embed_png <- function(path, dpi = NULL) {
  meta <- attr(png::readPNG(path, native = TRUE, info = TRUE), "info")
  if (!is.null(dpi)) meta$dpi <- rep(dpi, 2)

  knitr::asis_output(paste0(
    "<img src='", path, "'",
    " width=", round(meta$dim[1] / (meta$dpi[1] / 96)),
    " height=", round(meta$dim[2] / (meta$dpi[2] / 96)),
    " />"
  ))
}

knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
```

SelectorGadget is a JavaScript bookmarklet that allows you to interactively figure out what css selector you need to extract desired components from a page. 

## Installation

To install it, open this page in your browser, and then drag the following link to your bookmark bar: <a href="javascript:(function(){var%20s=document.createElement('div');s.innerHTML='Loading...';s.style.color='black';s.style.padding='20px';s.style.position='fixed';s.style.zIndex='9999';s.style.fontSize='3.0em';s.style.border='2px%20solid%20black';s.style.right='40px';s.style.top='40px';s.setAttribute('class','selector_gadget_loading');s.style.background='white';document.body.appendChild(s);s=document.createElement('script');s.setAttribute('type','text/javascript');s.setAttribute('src','https://dv0akt2986vzh.cloudfront.net/unstable/lib/selectorgadget.js');document.body.appendChild(s);})();">SelectorGadget</a>.

## Use

To use it, open the page 

1. Click on the element you want to select. SelectorGadget will make a first
   guess at what css selector you want. It's likely to be bad since it only
   has one example to learn from, but it's a start. Elements that match the
   selector will be highlighted in yellow.

2. Click on elements that shouldn't be selected. They will turn red.
   Click on elements that *should* be selected. They will turn green.

3. Iterate until only the elements you want are selected.  SelectorGadget
   isn't perfect and sometimes won't be able to find a useful css selector. 
   Sometimes starting from a different element helps.

For example, imagine we want to find the actors listed on an IMDB movie page, e.g. [The Lego Movie](http://www.imdb.com/title/tt1490017/).

1.  Navigate to the page and scroll to the actors list.

    ```{r, echo = FALSE}
    embed_png("selectorgadget-1.png")
    ```

2. Click on the SelectorGadget link in the bookmarks. The SelectorGadget
   console will appear at the bottom of the screen, and element currently
   under the mouse will be highlighted in orange.
   
    ```{r, echo = FALSE}
    embed_png("selectorgadget-2.png")
    ```

3.  Click on the element you want to select (the name of an actor). The 
    element you selected will be highlighted in green. SelectorGadget guesses
    which css selector you want (`.itemprop` in this case), and highlights
    all matches in yellow.

    ```{r, echo = FALSE}
    embed_png("selectorgadget-3.png")
    ```

4.  Scroll around the document to find elements that you don't want to match 
    and click on them. For example, we don't to match the title of the movie,
    so we click on it and it turns red. The css selector updates to 
    `#titleCast .itemprop`.

    ```{r, echo = FALSE}
    embed_png("selectorgadget-4.png")
    embed_png("selectorgadget-5.png")
    ```

Once we've determined the css selector, we can use it in R to extract the values we want:

```{r}
library(rvest)
lego_url <- "http://www.imdb.com/title/tt1490017/"
html <- read_html(lego_url)
cast <- html_nodes(html, ".primary_photo+ td a")
length(cast)
cast[1:2]
```
Finally, we can extract the text from the selected HTML nodes.

```{r}
html_text(cast, trim = TRUE)
```

Let's say we're also interested in extracting the links to the actors' pages. We can access html attributes of the selected nodes using `html_attrs()`.   

```{r}
cast_attrs <- html_attrs(cast)

length(cast_attrs)
cast_attrs[1:2]
```
As we can see there's only one attribute called `href` which contains relative url to the actor's page. We can extract it using `html_attr()`, indicating the name of the attribute of interest. Relative urls can be turned to absolute urls using `url_absolute()`.

```{r}
cast_rel_urls <- html_attr(cast, "href")
length(cast_rel_urls)
cast_rel_urls[1:2]

cast_abs_urls <- html_attr(cast, "href") %>% 
  url_absolute(lego_url)
cast_abs_urls[1:2]
```

